---
title: "Field Old and Young Leaves Lipid Analysis"
author: "Kenia E. Segura Aba"
date: "4/30/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align="center", fig.height=5, message=FALSE, warning=FALSE, fig.width=8, tidy.opts=list(width.cutoff=60),tidy=TRUE)
```
This file was run in R version 3.5.3. The packages used are `tidyverse` version 1.3.0, `readr` version 1.3.1, `RRPP` version 0.4.2.9000, `mixOmics` version 6.6.2, and `labdsv` version 2.0-1. This file must be in the same directory as the Box sync folder in order to run. The following analysis of lipid metabolites was conducted using a split-plot analysis of variance (ANOVA) of Young and Old *P. virgatum* leaves using residual randomization permutation procedure (`RRPP`). Patterns in metabolite classification were visualized using `mixOmics` for principle component analysis (PCA) and partial least squares discriminant analysis (PLS-DA). Dufrene-Legendre indicator analysis was performed to identify specific metabolites indicative of plant response to water treatment and fungal treatment (`labdsv`).

1. Load necessary packages
```{r}
library(tidyverse)
library(readr)
library(RRPP)
library(mixOmics)
library(labdsv)
```

# Lipids (Neg)

## RRPP

2. Define dependent variable matrix and class matrix.
```{r}
path <- "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/"

O_L_neg <- read_tsv(paste(path,"XCMS Online Results/O_L_Neg/XCMS.annotated.Report_1394368.tsv", sep=""))

Y_L_neg <- read_tsv(paste(path,"XCMS Online Results/Y_L_Neg/XCMS.annotated.Report_1394408.tsv", sep=""))

# dependent variable: metabolite intensities
Y_old <- O_L_neg[,c(2,12:54)] %>% data.frame(row.names=1) %>% t %>% data.frame()
scaled_Y_old <- scale(Y_old)

Y_young <- Y_L_neg[,c(2,12:54)] %>% data.frame(row.names=1) %>% t %>% data.frame()
scaled_Y_young <- scale(Y_young)

# class: sample factors
class <- read.csv(paste(path,"XCMS Online Results/class.csv", sep=""), header = T, row.names = 1)
```

3. Define and run multivariate regression models, then print out the results.
```{r}
O_LMneg <- lm.rrpp(scaled_Y_old ~ Block * Water * Fungus, data = class, SS.type = "III", print.progress = F); summary(O_LMneg)
Y_LMneg <- lm.rrpp(scaled_Y_young ~ Block * Water * Fungus, data = class, SS.type = "III", print.progress = F); summary(Y_LMneg)
```

4. Examine RRPP plots to check for assumptions.
```{r}
## Old Leaves
# residuals vs fitted values (homoscedasticity check)
Odiagnostics <- plot(O_LMneg, type = "diagnostics")

# linear regression plot
Oregression <- plot(O_LMneg, type = "regression", predictor = class$Fungus, reg.type = "RegScore")

# pca plot
Opcplot <- plot(O_LMneg, type = "PC", pch = 19, col = interaction(class$Water, class$Fungus))


## Young Leaves
# residuals vs fitted values (homoscedasticity check)
Ydiagnostics <- plot(Y_LMneg, type = "diagnostics")

# linear regression plot
Yregression <- plot(Y_LMneg, type = "regression", predictor = class$Fungus, reg.type = "RegScore")

# pca plot
Ypcplot <- plot(Y_LMneg, type = "PC", pch = 19, col = interaction(class$Water, class$Fungus))
```

5. Perform an RRPP ANOVA and print results.
```{r}
## Old Leaves
OnegANOVA <- anova(O_LMneg, effect.type = "F", error = c("Residuals", "Block:Water", "Block:Water:Fungus", "Residuals", "Block:Water:Fungus", "Block:Water:Fungus", "Residuals")) ; summary(OnegANOVA, formula = T)


## Young Leaves
YnegANOVA <- anova(Y_LMneg, effect.type = "F", error = c("Residuals", "Block:Water", "Block:Water:Fungus", "Residuals", "Block:Water:Fungus", "Block:Water:Fungus", "Residuals")) ; summary(YnegANOVA, formula = T)
```

6. Test lm.rrpp model coefficients. "d" is the amount of change in a variable for the coefficient indicated.
```{r}
## Old Leaves
# test model coefficients
Onegcoef <- coef(O_LMneg, test = T) ; summary(Onegcoef)

## Young Leaves
# test model coefficients
Ynegcoef <- coef(Y_LMneg, test = T) ; summary(Ynegcoef)
```
Fungus has the largest effect on the model for young leaves. The standard is the mean for High water treatment. Block has the largest effect on the model for old leaves. Fungus coming in second.

7. Compute predicted values from the lm.rrpp model fit using bootstrapped residuals to generate confidence intervals (precision of group mean estimates). 
```{r}
O_pred <- predict(O_LMneg) ; plot(O_pred, PC = T, ellipse = T)

Y_pred <- predict(Y_LMneg) ; plot(Y_pred, PC = T, ellipse = T)
```

8. Test pairwise differences between least squares means. Similar to tukeyHSD function in the r stats package. The pairwise function will generate tables with confidence intervals and p‐values for the pairwise statistic, Euclidean distance between least‐squares means.
```{r}
## Old Leaves
# pairwise differences of water 
Onegpw <- pairwise(O_LMneg, groups = class$Fungus); summary(Onegpw, confidence = 0.95, stat.table = T)

## Young Leaves
# pairwise differences of water
Ynegpw <- pairwise(Y_LMneg, groups = class$Fungus); summary(Ynegpw, confidence = 0.95, stat.table = T)
```


## PCA

8. Identify the major source of variation in data and determine if the variation is sourced from experimental bias or biological conditions.
```{r}
# Old Leaf Lipids (Neg)
# tune how many components to use
tune.pca(scaled_Y_old)

pca.res <- mixOmics::pca(scaled_Y_old, ncomp = 4, scale = F)

# plot pca
plotIndiv(pca.res, group = class$Fungus, ind.names = F, pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Old Leaf Lipids (Neg) PCA")

# Look at variable coefficients in each component with the loading vectors
# The absolute value of loading vectors represent the importance of each
# variable to define each PC
plotLoadings(pca.res, ndisplay = 50)

# Young Leaf Lipids (Neg)
# tune how many components to use
tune.pca(scaled_Y_young)

pca.res <- mixOmics::pca(scaled_Y_young, ncomp = 3, scale = F)

# plot pca
plotIndiv(pca.res, group = class$Fungus, ind.names = F ,pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Young Leaf Lipids (Neg) PCA")

# Look at variable coefficients in each component with the loading vectors
# The absolute value of loading vectors represent the importance of each
# variable to define each PC
plotLoadings(pca.res, ndisplay = 50)
```

9. Identify the variables that contribute to the explained variance.
```{r}
# Old Leaf
spca.res <- mixOmics::spca(scaled_Y_old, ncomp = 4, keepX = c(100,100,10,10))

# plot spca
plotIndiv(spca.res, group = class$Fungus, ind.names = F, pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Old Leaf Lipids (Neg) sPCA")

# variables contributing to each component
plotVar(spca.res, cex = 1)
selectVar(spca.res, comp = 1)$value # view loading value of each metabolite

# plot loadings for comp 1
plotLoadings(spca.res, ndisplay = 50)

# plot loadings for comp 2
plotLoadings(spca.res, comp=2, ndisplay = 50)


# Young Leaf
spca.res <- mixOmics::spca(scaled_Y_young, ncomp = 3, keepX = c(100,100,10))

# plot spca
plotIndiv(spca.res, group = class$Fungus, ind.names = F, pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Young Leaf Lipids (Neg) sPCA")

# variables contributing to each component
plotVar(spca.res, cex = 1)
selectVar(spca.res, comp = 1)$value # view loading value of each metabolite

# plot loadings for comp 1
plotLoadings(spca.res, ndisplay = 50)

# plot loadings for comp 2
plotLoadings(spca.res, comp=2, ndisplay = 50)
```

## PLS-DA

10. Classify samples into known groups and predict the class of new samples. 
```{r}
# Old Leaf
old.splsda <- mixOmics::splsda(scaled_Y_old, class$Fungus, keepX = c(100,100))

# plot pls-da
plotIndiv(old.splsda, ind.names = F, legend = T, title = "Old Leaf Lipids (Neg) PLS-DA", legend.title = "Fungus", ellipse = T)

# plot and select the variables
plotVar(old.splsda)
selectVar(old.splsda, comp=1)
plotLoadings(old.splsda, contrib = 'max', method = 'mean', ndisplay = 50)

# Young Leaf
young.splsda <- mixOmics::splsda(scaled_Y_young, class$Fungus, keepX = c(100,100))

# plot pls-da
plotIndiv(young.splsda, ind.names = F, legend = T, title = "Young Leaf Lipids (Neg) PLS-DA", legend.title = "Fungus", ellipse = T)

# plot and select the variables
plotVar(young.splsda)
selectVar(young.splsda, comp=1)
plotLoadings(young.splsda, contrib = 'max', method = 'mean', ndisplay = 50)
```

## Heatmaps of Averaged Data

11. Create averaged metabolite matrices and rerun PLS-DA to create a heatmap.
```{r}
## Young Leaves
av_Y_young <- aggregate(Y_young, by = list(class$Water, class$Fungus), FUN = "mean", simplify = T, data = class)
av.young.plsda <- mixOmics::plsda(av_Y_young[,3:1605], av_Y_young$Group.2) # fungus

# heatmap
youngcim <- cim(av.young.plsda, title = "Young Leaf Lipids (neg) Averaged Over Fungi", col.names = F, xlab = "Lipids", save = 'png', name.save = "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/Lipids Statistics/young_fungi_avlneg_hm.png") # by fungal treatment
```

## Indicator Analysis

12. Identify indicator metabolites characteristic of each treatment using Dufrene-Legendre Indicator Analysis.
```{r}
# Old Leaf
indicator_FungusO <- indval(Y_old, clustering = class$Fungus, numitr = 999, type = "long")
summary(indicator_FungusO)

# Young Leaf
indicator_Fungus <- indval(Y_young, clustering = class$Fungus, numitr = 999, type = "long")
summary(indicator_Fungus)
```

13. Disect indval object.
```{r}
Orelfrq <- indicator_FungusO$relfrq # relative frequency of species in classes
Orelabu <- indicator_FungusO$relabu # relative abundance of species in classes
Oindval <- indicator_FungusO$indval # the indicator value for each species
Omaxcls <- data.frame(indicator_FungusO$maxcls) # the class each species has max indicator value for
Oindcls <- data.frame(indicator_FungusO$indcls) # the indicator value for each species to its max class
Opval <- data.frame(indicator_FungusO$pval) # the probability of obtaining as high an indicator value as observed over the specified iterations

Yrelfrq <- indicator_Fungus$relfrq # relative frequency of species in classes
Yrelabu <- indicator_Fungus$relabu # relative abundance of species in classes
Yindval <- indicator_Fungus$indval # the indicator value for each species
Ymaxcls <- data.frame(indicator_Fungus$maxcls) # the class each species has max indicator value for
Yindcls <- data.frame(indicator_Fungus$indcls) # the indicator value for each species to its max class
Ypval <- data.frame(indicator_Fungus$pval) # the probability of obtaining as high an indicator value as observed over the specified iterations
```

14. Export results to a csv file.
```{r}
write.csv(cbind(Orelfrq, Orelabu, Oindval, Omaxcls, Oindcls, Opval), "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/Lipids Statistics/Indicator_Analys_oLneg_Fungus.csv")
write.csv(cbind(Yrelfrq, Yrelabu, Yindval, Ymaxcls, Yindcls, Ypval), "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/Lipids Statistics/Indicator_Analys_yLneg_Fungus.csv")
```

---

# Lipids (Pos)

## RRPP

2. Define dependent variable matrix and class matrix.
```{r}
path <- "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/"

O_L_pos <- read_tsv(paste(path,"XCMS Online Results/O_L_Pos/XCMS.annotated.Report_1394374.tsv", sep=""))

Y_L_pos <- read_tsv(paste(path,"XCMS Online Results/Y_L_Pos/XCMS.annotated.Report_1394370.tsv", sep=""))

# dependent variable: metabolite intensities
Y_old <- O_L_pos[,c(2,12:54)] %>% data.frame(row.names=1) %>% t %>% data.frame()
scaled_Y_old <- scale(Y_old)

Y_young <- Y_L_pos[,c(2,12:54)] %>% data.frame(row.names=1) %>% t %>% data.frame()
scaled_Y_young <- scale(Y_young)

# class: sample factors
class <- read.csv(paste(path,"XCMS Online Results/class.csv", sep=""), header = T, row.names = 1)
```

3. Define and run multivariate regression models, then print out the results.
```{r}
O_LMpos <- lm.rrpp(scaled_Y_old ~ Block * Water * Fungus, data = class, SS.type = "III", print.progress = F); summary(O_LMpos)
Y_LMpos <- lm.rrpp(scaled_Y_young ~ Block * Water * Fungus, data = class, SS.type = "III", print.progress = F); summary(Y_LMpos)
```

4. Examine RRPP plots to check for assumptions.
```{r}
## Old Leaves
# residuals vs fitted values (homoscedasticity check)
Odiagnostics <- plot(O_LMpos, type = "diagnostics")

# linear regression plot
Oregression <- plot(O_LMpos, type = "regression", predictor = class$Fungus, reg.type = "RegScore")

# pca plot
Opcplot <- plot(O_LMpos, type = "PC", pch = 19, col = interaction(class$Water, class$Fungus))


## Young Leaves
# residuals vs fitted values (homoscedasticity check)
Ydiagnostics <- plot(Y_LMpos, type = "diagnostics")

# linear regression plot
Yregression <- plot(Y_LMpos, type = "regression", predictor = class$Fungus, reg.type = "RegScore")

# pca plot
Ypcplot <- plot(Y_LMpos, type = "PC", pch = 19, col = interaction(class$Water, class$Fungus))
```

5. Perform an RRPP ANOVA and print results.
```{r}
## Old Leaves
OposANOVA <- anova(O_LMpos, effect.type = "F", error = c("Residuals", "Block:Water", "Block:Water:Fungus", "Residuals", "Block:Water:Fungus", "Block:Water:Fungus", "Residuals")) ; summary(OposANOVA, formula = T)


## Young Leaves
YposANOVA <- anova(Y_LMpos, effect.type = "F", error = c("Residuals", "Block:Water", "Block:Water:Fungus", "Residuals", "Block:Water:Fungus", "Block:Water:Fungus", "Residuals")) ; summary(YposANOVA, formula = T)
```

6. Test lm.rrpp model coefficients. "d" is the amount of change in a variable for the coefficient indicated.
```{r}
## Old Leaves
# test model coefficients
Oposcoef <- coef(O_LMpos, test = T) ; summary(Oposcoef)

## Young Leaves
# test model coefficients
Yposcoef <- coef(Y_LMpos, test = T) ; summary(Yposcoef)
```
Block and Fungus have the largest effect on the model for young leaves, but not their interaction. The standard is the mean for High water treatment.

7. Compute predicted values from the lm.rrpp model fit using bootstrapped residuals to generate confidence intervals (precision of group mean estimates). 
```{r}
O_pred <- predict(O_LMpos) ; plot(O_pred, PC = T, ellipse = T)

Y_pred <- predict(Y_LMpos) ; plot(Y_pred, PC = T, ellipse = T)
```

8. Test pairwise differences between least squares means. Similar to tukeyHSD function in the r stats package. The pairwise function will generate tables with confidence intervals and p‐values for the pairwise statistic, Euclidean distance between least‐squares means.
```{r}
## Old Leaves
# pairwise differences of fungus 
Opospw <- pairwise(O_LMpos, groups = class$Fungus); summary(Opospw, confidence = 0.95, stat.table = T)

## Young Leaves
# pairwise differences of fungus
Ypospw <- pairwise(Y_LMpos, groups = class$Fungus); summary(Ypospw, confidence = 0.95, stat.table = T)
```


## PCA

8. Identify the major source of variation in data and determine if the variation is sourced from experimental bias or biological conditions.
```{r}
# Old Leaf Lipids (Pos)
# tune how many components to use
tune.pca(scaled_Y_old)

pca.res <- mixOmics::pca(scaled_Y_old, ncomp = 3, scale = F)

# plot pca
plotIndiv(pca.res, group = class$Fungus, ind.names = F, pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Old Leaf Lipids (Pos) PCA")

# Look at variable coefficients in each component with the loading vectors
# The absolute value of loading vectors represent the importance of each
# variable to define each PC
plotLoadings(pca.res, ndisplay = 50)

# Young Leaf Lipids (Pos)
# tune how many components to use
tune.pca(scaled_Y_young)

pca.res <- mixOmics::pca(scaled_Y_young, ncomp = 3, scale = F)

# plot pca
plotIndiv(pca.res, group = class$Fungus, ind.names = F ,pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Young Leaf Lipids (Pos) PCA")

# Look at variable coefficients in each component with the loading vectors
# The absolute value of loading vectors represent the importance of each
# variable to define each PC
plotLoadings(pca.res, ndisplay = 50)
```

9. Identify the variables that contribute to the explained variance.
```{r}
# Old Leaf
spca.res <- mixOmics::spca(scaled_Y_old, ncomp = 3, keepX = c(100,100,10))

# plot spca
plotIndiv(spca.res, group = class$Fungus, ind.names = F, pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Old Leaf Lipids (Pos) sPCA")

# variables contributing to each component
plotVar(spca.res, cex = 1)
selectVar(spca.res, comp = 1)$value # view loading value of each metabolite

# plot loadings for comp 1
plotLoadings(spca.res, ndisplay = 50)

# plot loadings for comp 2
plotLoadings(spca.res, comp=2, ndisplay = 50)


# Young Leaf
spca.res <- mixOmics::spca(scaled_Y_young, ncomp = 3, keepX = c(100,100,10))

# plot spca
plotIndiv(spca.res, group = class$Fungus, ind.names = F, pch = as.factor(class$Water), legend = T, legend.title = "Fungus", legend.title.pch = "Water", title = "Young Leaf Lipids (Pos) sPCA")

# variables contributing to each component
plotVar(spca.res, cex = 1)
selectVar(spca.res, comp = 1)$value # view loading value of each metabolite

# plot loadings for comp 1
plotLoadings(spca.res, ndisplay = 50)

# plot loadings for comp 2
plotLoadings(spca.res, comp=2, ndisplay = 50)
```

## PLS-DA

10. Classify samples into known groups and predict the class of new samples. 
```{r}
# Old Leaf
old.splsda <- mixOmics::splsda(scaled_Y_old, class$Fungus, keepX = c(100,100))

# plot pls-da
plotIndiv(old.splsda, ind.names = F, legend = T, title = "Old Leaf Lipids (Pos) PLS-DA", legend.title = "Fungus", ellipse = T)

# plot and select the variables
plotVar(old.splsda)
selectVar(old.splsda, comp=1)
plotLoadings(old.splsda, contrib = 'max', method = 'mean', ndisplay = 50)

# Young Leaf
young.splsda <- mixOmics::splsda(scaled_Y_young, class$Fungus, keepX = c(100,100))

# plot pls-da
plotIndiv(young.splsda, ind.names = F, legend = T, title = "Young Leaf Lipids (Pos) PLS-DA", legend.title = "Fungus", ellipse = T)

# plot and select the variables
plotVar(young.splsda)
selectVar(young.splsda, comp=1)
plotLoadings(young.splsda, contrib = 'max', method = 'mean', ndisplay = 50)
```

## Heatmaps of Averaged Data

11. Create averaged metabolite matrices and rerun PLS-DA to create a heatmap.
```{r}
## Young Leaves
av_Y_young <- aggregate(Y_young, by = list(class$Water, class$Fungus), FUN = "mean", simplify = T, data = class)
av.young.plsda <- mixOmics::plsda(av_Y_young[,3:1755], av_Y_young$Group.2) # fungus

# heatmap
youngcim <- cim(av.young.plsda, title = "Young Leaf Lipids (pos) Averaged Over Fungi", col.names = F, xlab = "Lipids", save = 'png', name.save = "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/Lipids Statistics/young_fungi_avlpos_hm.png") # by fungal treatment
```

## Indicator Analysis

12. Identify indicator metabolites characteristic of each treatment using Dufrene-Legendre Indicator Analysis.
```{r}
# Old Leaf
indicator_FungusO <- indval(Y_old, clustering = class$Fungus, numitr = 999, type = "long")
summary(indicator_FungusO)
# Young Leaf
indicator_Fungus <- indval(Y_young, clustering = class$Fungus, numitr = 999, type = "long")
summary(indicator_Fungus)
```

13. Disect indval object.
```{r}
Orelfrq <- indicator_FungusO$relfrq # relative frequency of species in classes
Orelabu <- indicator_FungusO$relabu # relative abundance of species in classes
Oindval <- indicator_FungusO$indval # the indicator value for each species
Omaxcls <- data.frame(indicator_FungusO$maxcls) # the class each species has max indicator value for
Oindcls <- data.frame(indicator_FungusO$indcls) # the indicator value for each species to its max class
Opval <- data.frame(indicator_FungusO$pval) # the probability of obtaining as high an indicator value as observed over the specified iterations

Yrelfrq <- indicator_Fungus$relfrq # relative frequency of species in classes
Yrelabu <- indicator_Fungus$relabu # relative abundance of species in classes
Yindval <- indicator_Fungus$indval # the indicator value for each species
Ymaxcls <- data.frame(indicator_Fungus$maxcls) # the class each species has max indicator value for
Yindcls <- data.frame(indicator_Fungus$indcls) # the indicator value for each species to its max class
Ypval <- data.frame(indicator_Fungus$pval) # the probability of obtaining as high an indicator value as observed over the specified iterations
```

14. Export results to a csv file.
```{r}
write.csv(cbind(Orelfrq, Orelabu, Oindval, Omaxcls, Oindcls, Opval), "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/Lipids Statistics/Indicator_Analys_oLpos_Fungus.csv")
write.csv(cbind(Yrelfrq, Yrelabu, Yindval, Ymaxcls, Yindcls, Ypval), "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Kenia_Thesis_Analysis/Lipids Statistics/Indicator_Analys_yLpos_Fungus.csv")
```


---

## References

Collyer, M.L., Adams, D.C. 2018. RRPP: An r package for fitting linear models to high-dimensional data using residual randomization. Methods in Ecology and Evolution. 9(7):1772-1779.

Dufrene, M. and Legendre, P. 1997. Species assemblages and indicator species: the need for a flexible asymmetrical approach. Ecol. Monogr. 67(3):345-366.

Rohart, F., Gautier, B., Singh, A., & Lê Cao, K. A. 2017. mixOmics: An R package for ‘omics feature selection and multiple data integration. PLoS computational biology, 13(11):e1005752.
