---
title: "Field Chlorophyll Analysis of Old and Young Leaves"
author: "Kenia E. Segura Aba"
date: "5/2/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align="center", fig.height=5, message=FALSE, warning=FALSE, fig.width=8, tidy.opts=list(width.cutoff=60),tidy=TRUE)
```
This tutorial uses R version 4.0.0, `RRPP` version 0.5.2, `tidyverse` version 1.3.0, and `kableExtra` version 1.1.0.


1. load necessary packages
```{r}
library(RRPP)
library(tidyverse)
library(kableExtra)
```

2. Upload data
```{r}
path <- "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Old + Young Chlorophyll/"

chldata <- read.csv(paste(path, "chlorophyll_data.csv", sep=""), header = T, row.names = 1)

# create a total chlorophyll column (a + b)
chldata$Total <- chldata$Chl.a + chldata$Chl.b
```

3. Calculate total chlorophyll content, summary statistics, and visualize data.
```{r}
# plot chl a
chla <- chldata %>% ggplot(aes(x=Fungus, y=Chl.a, fill=Water)) +  geom_boxplot() + theme_classic() + facet_wrap(~Age,dir="v")
print(chla + labs(title="Chlorophyll a Content in Leaves"))

# plot chl b
chlb <- chldata %>% ggplot(aes(x=Fungus, y=Chl.b, fill=Water)) + geom_boxplot() +  theme_classic() + facet_wrap(~Age,dir="v")
print(chlb + labs(title="Chlorophyll b Content in Leaves"))

# plot total chlorophyll
tot <- chldata %>% ggplot(aes(x=Fungus, y=Total, fill=Water)) + geom_boxplot() + theme_classic() + facet_wrap(~Age,dir="v")
print(tot + labs(title="Total Chlorophyll Content in Leaves"))

# plot chl b/a ratio
chlb.a <- chldata %>% ggplot(aes(x=Fungus, y=Chl.b.a.ratio, fill=Water)) +  geom_boxplot() + theme_classic() + facet_wrap(~Age,dir="v")
print(chlb.a + labs(title="Chlorophyll b/a Ratio"))

# summary statistics
stats <- chldata %>% group_by(Water, Fungus) %>% select(Chl.a) %>% summarise_each(funs(Min = min, Q1 = quantile(., 0.25), Median = median, Q3 = quantile(., 0.75), Max = max, Mean = mean, SD = sd))

stats2 <- chldata %>% group_by(Water, Fungus) %>% select(Chl.b) %>% summarise_each(funs(Min = min, Q1 = quantile(., 0.25), Median = median, Q3 = quantile(., 0.75), Max = max, Mean = mean, SD = sd))

stats3 <- chldata %>% group_by(Water, Fungus) %>% select(Chl.b.a.ratio) %>% summarise_each(funs(Min = min, Q1 = quantile(., 0.25), Median = median, Q3 = quantile(., 0.75), Max = max, Mean = mean, SD = sd)) 

stats4 <- chldata %>% group_by(Water, Fungus) %>% select(Total) %>% summarise_each(funs(Min = min, Q1 = quantile(., 0.25), Median = median, Q3 = quantile(., 0.75), Max = max, Mean = mean, SD = sd))


stats %>% kable() %>% kable_styling(full_width = F) %>% add_header_above(c(" " = 2, "Chl a" = 7)) %>% collapse_rows()
stats2 %>% kable() %>% kable_styling(full_width = F) %>% add_header_above(c(" " = 2, "Chl b" = 7)) %>% collapse_rows()
stats3 %>% kable() %>% kable_styling(full_width = F) %>% add_header_above(c(" " = 2, "Chl b/a" = 7)) %>% collapse_rows()
stats4 %>% kable() %>% kable_styling(full_width = F,) %>% add_header_above(c(" " = 2, "Total Chlorophyll" = 7)) %>% collapse_rows()

```

4. Run Linear models in RRPP for each fungal treatment and estimate model coefficients. "d" is the amount of change in a variable for the coefficient indicated.
```{r}
high <- chldata %>% filter(Water=="High") %>% droplevels.data.frame()
low <- chldata %>% filter(Water=="Low") %>% droplevels.data.frame()

# chl b/a ratio linear models and coefficient tests
highLM <- lm.rrpp(Chl.b.a.ratio ~ Block*Fungus*Age, data=high, SS.type="III", print.progress=F) ; summary(highLM)
coef(highLM, test = T)

lowLM <- lm.rrpp(Chl.b.a.ratio ~ Block*Fungus*Age, data=low, SS.type="III", print.progress=F) ; summary(lowLM)
coef(lowLM, test = T)
```

5. Run ANOVA in RRPP for each linear model.
```{r}
# chl b/a ratio anovas
highANOVA <- anova(highLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(highANOVA)

lowANOVA <- anova(lowLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(lowANOVA)
```

6. Test pairwise differences between least squares means. Similar to tukeyHSD function in the r stats package. The pairwise function will generate tables with confidence intervals and p‐values for the pairwise statistic, Euclidean distance between least‐squares means.
```{r}
# chl b/a ratio pairwise
highpw <- pairwise(highLM, groups = high$Fungus) ; summary(highpw, confidence = 0.95, stat.table = T)
highpw2 <- pairwise(highLM, groups = high$Age) ; summary(highpw2, confidence = 0.95, stat.table = T)

lowpw <- pairwise(lowLM, groups = low$Fungus) ; summary(lowpw, confidence = 0.95, stat.table = T)
lowpw2 <- pairwise(lowLM, groups = low$Age) ; summary(lowpw2, confidence = 0.95, stat.table = T)
```

7. Examine RRPP plots to check for assumptions.
```{r}
## chl b/a ratio
# residuals vs fitted values (homoscedasticity check)
hdiagnostics <- plot(highLM, type = "diagnostics")

# pca plot
hpcplot <- plot(highLM, type = "PC", pch = 19, col = interaction(high$Water, high$Fungus))

# residuals vs fitted values (homoscedasticity check)
ldiagnostics <- plot(lowLM, type = "diagnostics")

# pca plot
lpcplot <- plot(lowLM, type = "PC", pch = 19, col = interaction(low$Water, low$Fungus))
```

8. Repeat steps 4 - 7 for chlorophyll a and b
```{r}
## chl a
# linear models and coefficient tests
highLM <- lm.rrpp(Chl.a ~ Block*Fungus*Age, data=high, SS.type="III", print.progress=F) ; summary(highLM)
coef(highLM, test = T)

lowLM <- lm.rrpp(Chl.a ~ Block*Fungus*Age, data=low, SS.type="III", print.progress=F) ; summary(lowLM)
coef(lowLM, test = T)

# ratio anovas
highANOVA <- anova(highLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(highANOVA)

lowANOVA <- anova(lowLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(lowANOVA)

# pairwise
highpw <- pairwise(highLM, groups = high$Fungus) ; summary(highpw, confidence = 0.95, stat.table = T)
highpw2 <- pairwise(highLM, groups = high$Age) ; summary(highpw2, confidence = 0.95, stat.table = T)

lowpw <- pairwise(lowLM, groups = low$Fungus) ; summary(lowpw, confidence = 0.95, stat.table = T)
lowpw2 <- pairwise(lowLM, groups = low$Age) ; summary(lowpw2, confidence = 0.95, stat.table = T)

# residuals vs fitted values (homoscedasticity check)
hdiagnostics <- plot(highLM, type = "diagnostics")

# pca plot
hpcplot <- plot(highLM, type = "PC", pch = 19, col = interaction(high$Water, high$Fungus))

# residuals vs fitted values (homoscedasticity check)
ldiagnostics <- plot(lowLM, type = "diagnostics")

# pca plot
lpcplot <- plot(lowLM, type = "PC", pch = 19, col = interaction(low$Water, low$Fungus))



## chl b
# linear models and coefficient tests
highLM <- lm.rrpp(Chl.b ~ Block*Fungus*Age, data=high, SS.type="III", print.progress=F) ; summary(highLM)
coef(highLM, test = T)

lowLM <- lm.rrpp(Chl.b ~ Block*Fungus*Age, data=low, SS.type="III", print.progress=F) ; summary(lowLM)
coef(lowLM, test = T)

# ratio anovas
highANOVA <- anova(highLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(highANOVA)

lowANOVA <- anova(lowLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(lowANOVA)

# pairwise
highpw <- pairwise(highLM, groups = high$Fungus) ; summary(highpw, confidence = 0.95, stat.table = T)
highpw2 <- pairwise(highLM, groups = high$Age) ; summary(highpw2, confidence = 0.95, stat.table = T)

lowpw <- pairwise(lowLM, groups = low$Fungus) ; summary(lowpw, confidence = 0.95, stat.table = T)
lowpw2 <- pairwise(lowLM, groups = low$Age) ; summary(lowpw2, confidence = 0.95, stat.table = T)

# residuals vs fitted values (homoscedasticity check)
hdiagnostics <- plot(highLM, type = "diagnostics")

# pca plot
hpcplot <- plot(highLM, type = "PC", pch = 19, col = interaction(high$Water, high$Fungus))

# residuals vs fitted values (homoscedasticity check)
ldiagnostics <- plot(lowLM, type = "diagnostics")

# pca plot
lpcplot <- plot(lowLM, type = "PC", pch = 19, col = interaction(low$Water, low$Fungus))



## Total
# linear models and coefficient tests
highLM <- lm.rrpp(Total ~ Block*Fungus*Age, data=high, SS.type="III", print.progress=F) ; summary(highLM)
coef(highLM, test = T)

lowLM <- lm.rrpp(Total ~ Block*Fungus*Age, data=low, SS.type="III", print.progress=F) ; summary(lowLM)
coef(lowLM, test = T)

# ratio anovas
highANOVA <- anova(highLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(highANOVA)

lowANOVA <- anova(lowLM, effect.type = "F", error = c("Residuals", "Block:Fungus", "Block:Fungus:Age", "Residuals", "Block:Fungus:Age", "Block:Fungus:Age", "Residuals")) ; summary(lowANOVA)

# pairwise
highpw <- pairwise(highLM, groups = high$Fungus) ; summary(highpw, confidence = 0.95, stat.table = T)
highpw2 <- pairwise(highLM, groups = high$Age) ; summary(highpw2, confidence = 0.95, stat.table = T)

lowpw <- pairwise(lowLM, groups = low$Fungus) ; summary(lowpw, confidence = 0.95, stat.table = T)
lowpw2 <- pairwise(lowLM, groups = low$Age) ; summary(lowpw2, confidence = 0.95, stat.table = T)

# residuals vs fitted values (homoscedasticity check)
hdiagnostics <- plot(highLM, type = "diagnostics")

# pca plot
hpcplot <- plot(highLM, type = "PC", pch = 19, col = interaction(high$Water, high$Fungus))

# residuals vs fitted values (homoscedasticity check)
ldiagnostics <- plot(lowLM, type = "diagnostics")

# pca plot
lpcplot <- plot(lowLM, type = "PC", pch = 19, col = interaction(low$Water, low$Fungus))
```

9. Indicator Analysis
```{r}
library(labdsv)
## High water
indicator_high <- indval(high[,5:8], clustering = high$Fungus, numitr = 999, type = "long")

write.csv(cbind(indicator_high$relfrq, indicator_high$relabu, indicator_high$indval, data.frame(indicator_high$maxcls), data.frame(indicator_high$indcls), data.frame(indicator_high$pval)), "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Old + Young Chlorophyll/Indicator_Analys_High_Fungus.csv")

## Low water
indicator_low <- indval(low[,5:8], clustering = low$Fungus, numitr = 999, type = "long")

write.csv(cbind(indicator_low$relfrq, indicator_low$relabu, indicator_low$indval, data.frame(indicator_low$maxcls), data.frame(indicator_low$indcls), data.frame(indicator_low$pval)), "~/Box/Summer 2018 TX Endo Field Samples and Analysis/Statistics/Old + Young Chlorophyll/Indicator_Analys_Low_Fungus.csv")
```

